---
title: Manage the Flow of Execution in R by Using Conditional Statements
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Requirements


## Evaluate a condition by using an if statement
```{r Evaluate a condition by using an if statement, include=T, error=F, warning=F}






```

## Evaluate a condition in an if statement by using the any() function
```{r Evaluate a condition in an if statement by using the any() function, include=T, error=F, warning=F}






```

## Evaluate a condition in an if statement by using the all() function
```{r Evaluate a condition in an if statement by using the all() function, include=T, error=F, warning=F}






```

## Control the flow of execution by using an if...else statement
```{r Control the flow of execution by using an if...else statement, include=T, error=F, warning=F}






```
## Evaluate a condition by using the ifelse() function
```{r Evaluate a condition by using the ifelse() function, include=T, error=F, warning=F}






```

## Evaluate multiple conditions by using a nested if statement
```{r Evaluate multiple conditions by using a nested if statement, include=T, error=F, warning=F}






```

## Control the flow of execution by using a switch statement
```{r Control the flow of execution by using a switch statement, include=T, error=F, warning=F}






```






# The switch statement







```
















