---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-011-Can you implement Data Frames and Factors in R? [Advanced]
## R Data Frame & Factor Resources
### Notes: R Data Frames
- Resources:
  - [data.frame()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/data.frame) function used to create R data frame objects
    - stringsAsFactors: default = TRUE; Sets character vectors to factors if TRUE
  - Data Frame Subsetting:
    - [How to Subset a Data Frame in R (4 Examples)](https://www.statology.org/subset-data-frame-in-r/)
    - (How To Subset An R Data Frame – Practical Examples)[https://www.programmingr.com/examples/r-dataframe/subset-an-r-data-frame/]
  - Data Frame Operations:
    - [merge()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/merge) function to merge/join two data frames

### Notes: R Factors
- Factors are used for withing with *categorical* or *nominal* data. This is type of data is nonnumerical and categorical in nature. 
  - **Categorica/Nominal** Data:
    - Non-numerical
    - Not continuous such as temperatures, heights, and weights
    - In this type of data an element or data point belongs to a distinct category
    - **Levels** are the distinct count of elements within a category(**factor**)
  - Examples include: 
    - Gender: Male|Female
    - Party Affiliation: Republican|Democrat|Independent
    - Marital Status: married|single|divorced|widowed
- Factors and R Table objects are frequently used together. 
- The R data frame default behavior is to make string data types factors. Set **stringAsFactor = FALSE** in the **data.frame()** method to diable this feature.
- Factors have *levesl* which is a disctinct value count

### Related Links
- [Factors in R](https://www.stat.berkeley.edu/~s133/factors.html) an article from UC Berkeley
- [factor()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/factor) function used to create factors in R
- [levels()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/levels) function for level attributes
- [table()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/table) function for cross tabulation and table creation

## Creeate R data frame objects
```{r Create R data frame object, include = T, error=F,warning=F}





```
## Join two data frames
```{r Join two data frames, include= T, warning=F, error=F}






```

## Select and query data from an R data frame
```{r Select and query data from an R data frame, include= T, error=F, warning=F}



```

## Modify an R data frame row
```{r Modify an R data frame row, include=T, error=F, warning=F}





```







