---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-014-Create Functions in R [Guided]

### Notes: R Functions
- R Functions are "first class" objects
- User defined R Functions are listed in the Environments pane
- Functions and accept optional arguments
- R supports function nesting which means functions can be declared within functions
- By default, a user defined R function will return the last evaluated expression 
- Defining a function
  - Created using the function() method
  - Syntax: No parameters
    ```
      f1 <- function(){
        # Do something
        # Do another thing
        # Do more things
      
      }
    ```
  - Syntax: With parameters
    ```
      # Must be called using parameter order (arguement position) or by arguement name
      f1 <- function(arg1, arg2){
        # Do something
        # Do another thing
        # Do more things
      
      }
      
      # Specifying default parameter values
      f2 <- function(arg1 = TRUE, arg2 = 'some value'){
      
        # Do something
        # Do another thing
        # Do more things
        
        # Maybe call return()
      
      }
      
      # Using ellipses
      f3 <- function(arg1 = TRUE, arg2 = 'some value', arg3, arg4 = 5, ...){
      
        # Do something
        # Do another thing
        # Do more things
        
        # Maybe call return()
      
      }
      
    ```
- Returning values:
  - By default R will implicitly return NULL or the last lexical scoped variable declared
  - In R a function has exactly one return value
    - Return a data structure such as a List if multiple values need to be returned
  - Syntax:
    ```
        f1 <- function(arg1, arg2){
          # Do something
          # Do another thing
          # Do more things
        
        return(#output a single value)
      
      }
    ```
### Important Resources:
- [function()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/function) for creating user defined functions
- [return()](https://faculty.nps.edu/sebuttre/home/R/functions.html) used to return *a* value
- [missing()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/missing) function used to test and check to see if functions arguements are missing
- [...](https://faculty.nps.edu/sebuttre/home/R/functions.html) ellipses are special characters used to mean "anything else." They are usually placed last in the parameter list
- [body()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/body) function used to get or set a function's body
- [formals()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/formals) function to get or set the formal arguements

### Related Resources: Usefull R functions
- [print()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/print) function to print values
- [format()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/format) function to encode output formats
- [cat()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cat) function to output and concatenate strings

## Create a function in R
```{r Create a function in R, include= T, warning=F, error=F}
# function definition
f_demo1 <- function(){
  cat('The value of pi is', pi)
}

# function call/function execution
f_demo1()

```
## Declare a variable inside a function
```{r Declare a variable inside a function, include=T, warning=F, error=F}
# function definition
  f_demo2 <- function(){
    
      var1 = "This is var1!"
      cat(var1)
    
  }
# function call/function execution
f_demo2()

```
## Implement function parameters
```{r Implement function parameters, include=T, warning=F,error=F}

  f_demo3 <- function(arg1) {
    var1 <- arg1
    cat('You passed', var1, 'to f_demo3')
  }

f_demo3('Greetings')
f_demo3(arg1 = "This is fun")


```

## Create a default function parameter
```{r Create a default function parameter, include=T, warning=F, error=F}

f_demo4 <- function(arg1, arg2 = "I'm a default value"){
  cat("arg1 =", arg1, "\narg2 =", arg2)
}

f_demo4("arg1 value", "arg2 value")

f_demo4("arg1 value")

```


## Print a function's body and formal argument list
```{r Print a function\'s body and formal argument list, include = T, error = F, warning = F }

body(f_demo4)

formals(f_demo4)


```

## Return an implicit and explicit value from a function
```{r Return an implic value, include=T, error=F, warning=T}

f_demo6 <- function(){
  
  var1 <- 78L
  var2 <- TRUE
  var3 <- "I'm var!"
  
}

var1 <- f_demo6()
var1


f_demo7 <- function(){
  
  var1 <- 78L
  var2 <- TRUE
  var3 <- "I'm var!"
  
  return(var1)
  
}


f_demo7()



```

