---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-013-Execute Flow Control Statements in R [Guided]

## Notes: Flow Control Statements
- Looping Methods
  - for loop: iterates the loop element-by-element
  - while loop: repeat code until a specified condition evaluates to FALSE
- for Loops
  - When to stop loop execution in known upfront. 
    - Ex:
      - Loop over the number vector elements
      - Loop over the rows in a data frame
  - loopindex: an individual element value within the loopvector
  - loopvector: an iterable object such as a vector
  - Syntax:
    ```
    for(loopindex in loopvector) {
      # code to loop and repeat
      
    }
    ```
- while loops:
  - Used when there is not a built in stop mechanism such as the number of vector elements.
    - The programmer must write their own counting logic determining when to stop loop execution.
  - Syntax:
    ```
    while(condition) {
      # code to iterate and execute
    
    }
    ```

## Related Resources: Usefull R functions
- apply family functions
  - Performs implicit looping
    - Can be used in place of a for loop in certain cases
    - The apply() function takes a function and *applies* it to each *margin* of an array
    - The apply function takes at least three arguments;
      - X: specifies the iterable objet
      - MARGIN: An integer specifying to iterate over rows (1) or columns (2).
      - FUN: The function to execute against the margin
  - [apply()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/apply) function used to apply functions over array margins
    - Best for arrays and matrices
  - [tapply()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/tapply) function used to apply a function over a Ragged Array
  - [sapply()](https://r-lang.com/r-sapply/) function used to apply a function to all input elements
  - [lapply()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/lapply) function used to apply a function over a List or Vector
- [print()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/print) function used to print values 
- [cat()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cat) function used to concatenate and print values 
- [R Constants](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Constants.html) built in base and core R constants
- [names()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/names) function to to get or set the names of an object
- [cbind()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/cbind) function to combine an object by columns (or rbind() for rows).

## Implement an R for loop
```{r Implement an R for loop, include=TRUE, error=F, warning=F}
for(idx in 1:10) {
  
  print(idx)
  
}

for (idx in month.name){
  print(idx)
}

```
## Compare two for loops using idexing and values
```{r Compare two for loops using indexing and values, include=T, warning=F, error=F}
v_months <- month.abb

for(idx in v_months){
  print(idx)
}

for(idx in 1:length(v_months)){
  
  cat('Month number', idx, v_months[idx], '\n')
  
}

```


## Implement for loop nesting
```{r Implement for loop nesting, include=TRUE, error=F, warning=F}
v_names <- c('Jim', 'Lynn', 'Joyce', 'Paul')
v_majors <- c('Finance', 'Education', 'Computer Science', 'Economics')
v_GPAs <- c(3.50, 3.50, 2.75, 3.00)

lst_classRoster <- list(Names = v_names, Majors = v_majors, GPAs = v_GPAs)
lst_classRoster

for(idx_classRoster in names(lst_classRoster)) {
  
  if(idx_classRoster == 'GPAs')
  {
    for(idx in lst_classRoster[idx_classRoster]){
      print(idx)
    }
  }
}
```

## Implement an R while loop
```{r Implement an R while loop, include=T, error=F, warning=F}

my_counter <- 0 

while(my_counter <= 25) {
  print(my_counter)
  my_counter <- my_counter + 1
}

```

## Implement implicit looping using the apply function
```{r Implement implicit looping using the apply function, include=T, warning=F, error=F}

m_test <- matrix(1:25, nrow=5, ncol=5)
m_test

apply(X = m_test, MARGIN = 1, FUN = sum)

m_test <- cbind(m_test, apply(X = m_test, MARGIN =1, FUN = sum))
m_test <- rbind(m_test, apply(X = m_test, MARGIN = 2, FUN = sum))
m_test
```






