---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-010-Implement R Factors [Guided]
## R Factor Resources
### Notes: R Factors
- Factors are used for withing with *categorical* or *nominal* data. This is type of data is nonnumerical and categorical in nature. 
  - **Categorica/Nominal** Data:
    - Non-numerical
    - Not continuous such as temperatures, heights, and weights
    - In this type of data an element or data point belongs to a distinct category
    - **Levels** are the distinct count of elements within a category(**factor**)
  - Examples include: 
    - Gender: Male|Female
    - Party Affiliation: Republican|Democrat|Independent
    - Marital Status: married|single|divorced|widowed
- Factors and R Table objects are frequently used together. 
- The R data frame default behavior is to make string data types factors. Set **stringAsFactor = FALSE** in the **data.frame()** method to diable this feature.
- Factors have *levesl* which is a disctinct value count

### Related Links
- [Factors in R](https://www.stat.berkeley.edu/~s133/factors.html) an article from UC Berkeley
- [factor()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/factor) function used to create factors in R
- [levels()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/levels) function for level attributes
- [table()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/table) function for cross tabulation and table creation

## Implement a simple R factor
```{r Implment a simple R factor, include=T, error=F, warning=F}




```

## Examine an R factor
```{r Examine an R factor, include=T, error=F, warning=FALSE}




```

## Use the levels attribute and levels function
```{r Use the levels attribute and levels function, include=T, error=F, warning=F}





```

## Implement factors in a data frame
```{r Implement factors in a data frame, include= T, error= F, warning=F}




```




