---
title: "Skillbuilder Base and Core R Programming Fundamentals"
author: "User 1"
date: "12/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-007-Implement R Matrices  [Guided]

## The R Matrix Data Structure Object
### R Data Structures
- Vector
  - The simplest and basic R data structure.
  - A one dimensional array. Can be scalar having a single value
  - All elements of the same data type
- List 
  - An ordered collection of objects
  - Elements can be of mixed data types
- Matrix
  - A 2x2 arrary. Tabular/spreadsheet in nature
  - created usin the matrix(), cbind(), or bind() functions. 
- DataFrame - Discussed later in this series.
- Factor - Discussed later in this series

# Create an R Matrix
```{r Intro, include= T, warning=F, error=F}

m1 <- matrix(1:9, ncol = 3, nrow = 3)
m1
dim(m1)

m2 <- matrix(1:9, ncol = 3, nrow = 3, byrow = TRUE)
m2

m_classRoster <- matrix(c(1001, "John", "Smith", "A", 1002, "Jane", "Jackson", "A+", 1003, "Ralph", "Pederson", "B-"), ncol = 4, nrow = 4, byrow = TRUE)
m_classRoster
dim(m_classRoster)



```

# Implementing the R Class Roster Matrix
```{r Implement, include=T, error=T, warning=T}

colnames(m_classRoster) = c("StudentID", "FirstName", "LastName", "Grade")
m_classRoster
m_classRoster[c(1,2), c(2,3)]
m_classRoster[, "FirstName"]




```

# Modifying the R Class Roster Matrix
```{r modify, include=T, warning=F, error=F}

m_classRoster[1, 2] <- "George"
m_classRoster

mt_classRoster = t(m_classRoster)
mt_classRoster

mt_classRoster <- NULL


```

