---
title: "Skillbuilder Base and Core R Programming Fundamentals"
author: "User 1"
date: "12/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-007-Implement R Matrices  [Guided]

## The R Matrix Data Structure Object
### R Data Structures
- Vector
  - The simplest and basic R data structure.
  - A one dimensional array. Can be scalar having a single value
  - All elements of the same data type
- List 
  - An ordered collection of objects
  - Elements can be of mixed data types
- Matrix
  - A 2x2 arrary. Tabular/spreadsheet in nature
  - created usin the matrix(), cbind(), or bind() functions. 
- DataFrame - Discussed later in this series.
- Factor - Discussed later in this series

# Create an R Matrix
```{r Simple R Matrix, include= T, warning=F, error=F}



```

# Implementing the R Class Roster Matrix
```{r Create & Implement a Class Roster, include=T, error=T, warning=T}



```

# Modifying the R Class Roster Matrix
```{r Modify the Class Roster Matrix, include=T, warning=F, error=F}




```
