---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-010-Implement R Factors [Guided]
## R Factor Resources
### Notes: R Factors
- Factors are used for withing with *categorical* or *nominal* data. This is type of data is nonnumerical and categorical in nature. 
  - **Categorica/Nominal** Data:
    - Non-numerical
    - Not continuous such as temperatures, heights, and weights
    - In this type of data an element or data point belongs to a distinct category
    - **Levels** are the distinct count of elements within a category(**factor**)
  - Examples include: 
    - Gender: Male|Female
    - Party Affiliation: Republican|Democrat|Independent
    - Marital Status: married|single|divorced|widowed
- Factors and R Table objects are frequently used together. 
- The R data frame default behavior is to make string data types factors. Set **stringAsFactor = FALSE** in the **data.frame()** method to diable this feature.
- Factors have *levesl* which is a disctinct value count

### Related Links
- [Factors in R](https://www.stat.berkeley.edu/~s133/factors.html) an article from UC Berkeley
- [factor()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/factor) function used to create factors in R
- [levels()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/levels) function for level attributes
- [table()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/table) function for cross tabulation and table creation

## Implement a simple R factor
```{r Implment a simple R factor, include=T, error=F, warning=F}
v_marriagestatus <- c('married', 'married', 'married', 'single', 'single', 'divorced', 'divorced','married','married','widower')

f_marriagestatus <- factor(v_marriagestatus)

f_marriagestatus

str(f_marriagestatus)

```

## Examine an R factor
```{r Examine an R factor, include=T, error=F, warning=FALSE}
length(f_marriagestatus)
levels(f_marriagestatus)
table(f_marriagestatus)


```

## Use the levels attribute and levels function
```{r Use the levels attribute and levels function, include=T, error=F, warning=F}
v_movie_genres <- c('Horror', 'SciFi', 'Drama', 'Horror', 'Horror', 'Horror', 'Drama', 'SciFi', 'Fantasy', 'Thriller', 'Thriller', 'SciFi')
f_movie_genres <- factor(v_movie_genres, levels = c('Horror', 'SciFi', 'Comedy', 'Drama', 'Other'))
f_movie_genres

levels(f_movie_genres)
table(f_movie_genres)

v_partyaffiliations <- c('R', 'R', 'R', 'I', 'D', 'D','D', 'R', 'R', 'I', 'D')
f_partyaffiliations <- factor(v_partyaffiliations)
levels(f_partyaffiliations)

levels(f_partyaffiliations) <- c(1, 2, 3)
levels(f_partyaffiliations)
table(f_partyaffiliations)


```

## Implement factors in a data frame
```{r Implement factors in a data frame, include= T, error= F, warning=F}
v_userid <- c(1001, 1002, 1003, 1004, 1005)
v_names <- c('Fred Johnson', 'Laura Smith', 'Bob Jones', 'Tyra Dance', 'Jimmy Peters')
f_affiliations <- factor(c('R', 'R', 'D', 'R', 'D'))
df_users <- data.frame(UserID = v_userid, Name = v_names, Party = f_affiliations, stringsAsFactors = F)
str(df_users)

df_users$Name[f_affiliations == 'R']
df_users[f_affiliations == 'D', ]

```




