---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-009-Implement R Data Frame Fundamentals [Guided]

## The R Data Frame:
- A 2 dimensional spreadsheet-like data structure similar to a database table
   - Think rows X columns
- An important R construct used R data structure in data science and statistical analysis
- Can be thought of as a special kind of list:
  - Can store various data types such as numbers, text, and logical variables
  - **CRITICAL**: Data frame members must all be equal length vectors.

## Create and examine R data frame object
```{r create an R data frame, include=T, warning=F, error=T}

v__StudentID <- c(1001,1002,1003)
v_FirstName <- c('John', 'Linda', 'Tyra')
v_LastName <- c('Smith', 'Jones', 'Johnson')
v_Major <- c('Computer Sci', 'Data Sci', 'Finance')
v_GPA <- c(3.75, 3.65,3.75)

df_classRoster <- data.frame( StudentID = v__StudentID, FirstName = v_FirstName, LastName = v_LastName, Major = v_Major, GPA = v_GPA, stringsAsFactors = FALSE )

df_classRoster
str(df_classRoster)
dim(df_classRoster)
colnames(df_classRoster)


```
## Select and query data from an R data frame
```{r select and query, include=T, error = T, warring = T}

df_classRoster[2:3]
df_classRoster[c('FirstName', 'LastName')]
df_classRoster[df_classRoster['GPA'] >= 3.75, ]

mean( df_classRoster[["GPA"]])

```

## Modify an R data frame
```{r modify an R data frame, include= T, error= F , warning= F}
df_classRoster <- rbind(df_classRoster, c(1004, 'Fred', 'Brown','Psychology', 2.75))
df_classRoster <- rbind(df_classRoster, c(1005, 'Angie', 'Watson', 'Business', 3.00))
df_classRoster

gpa_sort <- order(df_classRoster$GPA, decreasing = TRUE)
df_classRoster <- df_classRoster[gpa_sort, ]
df_classRoster

df_classRoster <- df_classRoster[df_classRoster$GPA > 3.00, ]
df_classRoster

```

# Data Frame Resouces:

## Links:
- [data.frame()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/data.frame)
  - [R CODER](https://r-coder.com/dataframe-r/) article, "DataFrame in R"
  - [stringAsFactors = TRUE(default) | FALSE](https://developer.r-project.org/Blog/public/2020/02/16/stringsasfactors/index.html)
- [str()](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/str) function for data frame structure
- [dim()](https://r-lang.com/dim-in-r/) function for data frame dimensions rowsXcols
- [colnames()](https://rdrr.io/r/base/colnames.html) function to list column names
- [head()]() function ouput the first n number of rows
- [tail()]() function output the last n number of rows
- [summary()]() function output summary statistics
- [cbind()]() function used to add columns
- [rbind()]() function used to add rows
- [order()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/order) function used to sort and order data




