---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-009-Implement R Data Frame Fundamentals [Guided]

## The R Data Frame:
- A 2 dimensional spreadsheet-like data structure similar to a database table
   - Think rows X columns
- An important R construct used R data structure in data science and statistical analysis
- Can be thought of as a special kind of list:
  - Can store various data types such as numbers, text, and logical variables
  - **CRITICAL**: Data frame members must all be equal length vectors.

## Create and examine R data frame object
```{r create an R data frame, include=T, warning=F, error=T}






```
## Select and query data from an R data frame
```{r select and query, include=T, error = T, warring = T}




```

## Modify an R data frame
```{r modify an R data frame, include= T, error= F , warning= F}




```

# Data Frame Resouces:

## Links:
- [data.frame()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/data.frame)
  - [R CODER](https://r-coder.com/dataframe-r/) article, "DataFrame in R"
  - [stringAsFactors = TRUE(default) | FALSE](https://developer.r-project.org/Blog/public/2020/02/16/stringsasfactors/index.html)
- [str()](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/str) function for data frame structure
- [dim()](https://r-lang.com/dim-in-r/) function for data frame dimensions rowsXcols
- [colnames()](https://rdrr.io/r/base/colnames.html) function to list column names
- [head()]() function ouput the first n number of rows
- [tail()]() function output the last n number of rows
- [summary()]() function output summary statistics
- [cbind()]() function used to add columns
- [rbind()]() function used to add rows
- [order()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/order) function used to sort and order data



