---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-011-Can you implement Data Frames and Factors in R? [Advanced]
## R Data Frame & Factor Resources
### Notes: R Data Frames
- Resources:
  - [data.frame()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/data.frame) function used to create R data frame objects
    - stringsAsFactors: default = TRUE; Sets character vectors to factors if TRUE
  - Data Frame Subsetting:
    - [How to Subset a Data Frame in R (4 Examples)](https://www.statology.org/subset-data-frame-in-r/)
    - (How To Subset An R Data Frame – Practical Examples)[https://www.programmingr.com/examples/r-dataframe/subset-an-r-data-frame/]
  - Data Frame Operations:
    - [merge()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/merge) function to merge/join two data frames

### Notes: R Factors
- Factors are used for withing with *categorical* or *nominal* data. This is type of data is nonnumerical and categorical in nature. 
  - **Categorica/Nominal** Data:
    - Non-numerical
    - Not continuous such as temperatures, heights, and weights
    - In this type of data an element or data point belongs to a distinct category
    - **Levels** are the distinct count of elements within a category(**factor**)
  - Examples include: 
    - Gender: Male|Female
    - Party Affiliation: Republican|Democrat|Independent
    - Marital Status: married|single|divorced|widowed
- Factors and R Table objects are frequently used together. 
- The R data frame default behavior is to make string data types factors. Set **stringAsFactor = FALSE** in the **data.frame()** method to diable this feature.
- Factors have *levesl* which is a disctinct value count

### Related Links
- [Factors in R](https://www.stat.berkeley.edu/~s133/factors.html) an article from UC Berkeley
- [factor()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/factor) function used to create factors in R
- [levels()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/levels) function for level attributes
- [table()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/table) function for cross tabulation and table creation

## Creeate R data frame objects
```{r Create R data frame object, include = T, error=F,warning=F}
v_userIDs <- c(1001, 1002, 1003)
v_FirstNames <- c('Fred', 'Mary', 'Peter')
v_LastNames <- c('Johnson', 'Smith', 'Reeves')
v_emails <- c('fred.johnson@test.com', 'mary_smith@test.com', 'reevp@example.org')

df_Users1 <- data.frame(UserID = v_userIDs, FirstName = v_FirstNames, LastName = v_LastNames, Email = v_emails, stringsAsFactors = FALSE)
df_Users1
str(df_Users1)

v_maritalstatus <- c('married', 'married', 'divorced')
f_maritalstatus <- factor( v_maritalstatus, levels = c('married', 'divorced', 'widowed'))
df_Users2 <- data.frame(UserID = v_userIDs, MaritalStatus = f_maritalstatus, stringsAsFactors = FALSE)
df_Users2
str(df_Users2)



```
## Join two data frames
```{r Join two data frames, include= T, warning=F, error=F}

df_Users <- merge(df_Users1, df_Users2, by = 'UserID')
df_Users

df_Users <- rbind(df_Users,c(1004,'Henry','Jones','jonesh@example.edu','married'))
df_Users <- rbind(df_Users,c(1005,'Linda','Washington','washlin@test.org','divorced'))
df_Users <- rbind(df_Users,c(1006,'Laura','Fine','laura.fine@test.us','divorced'))
df_Users
dim(df_Users)
colnames(df_Users)


```

## Select and query data from an R data frame
```{r Select and query data from an R data frame, include= T, error=F, warning=F}
df_Users[ df_Users['MaritalStatus'] == 'married', ]

df_Users[order(df_Users$LastName, decreasing = FALSE), ]


```

## Modify an R data frame row
```{r Modify an R data frame row, include=T, error=F, warning=F}

df_Users[df_Users['UserID'] ==1005, ]
df_Users[df_Users$UserID == 1005, ]$LastName <- 'Brown'
df_Users[df_Users$UserID == 1005, ]$MaritalStatus <- 'married'
df_Users[order(df_Users$LastName, decreasing = FALSE), ]


```







