---
title: 'Skillable: Introduction to R & RStudio Desktop'
author: "User 1"
date: "12/27/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# RRD-012-Implement Conditional Statements using R [Guided]

### Notes: Conditional Statements
- if statement:
  - Checks the conditional for a **single** logical variable 
  - Basic syntax:
    ```
    if(condition) {
      ## Execute if the condition evaluates to TRUE
    }
    ```
  - Single-line if statement without curry braces {}
    ```
      if(condition) ## Execute if TRUE
    ```
      - The line after a single-line if statement is not considered apart of the if statement and will run unconditionally
      ```
      if(condition) ## Do something if TRUE
      ## ALWAYS Do something
      ```
  - Curly braces {} are optional if there is a single statement after the if condition
  - All three examples function identically without regard for syntax
    - Example: 
    ```
      if( var1 == 9) 'Var1 equals 9'
    
      if ( var1 == 9) { 'Do something if TRUE' }
      
      if (var1 == 9) {
        'Do something if TRUE'
      }
      
    ```
  - Curly braces {} are **recommend** to make code easier to read, debug, and understand
  
- if-else statement:
  - If statements **only** run when the condition is TRUE. 
  - Else statements run when the the condition is FALSE
  - Syntax:
  ```
  # Option 1:
  if(condition){
    do any code in here if condition is TRUE
  } else {
    do any code in here if condition is FALSE
  }
  
  # You can take liberty with white space and format as such:
  
  # Option 2:
  if(condition)
  {
    do any code in here if condition is TRUE
  } 
  else 
  {
    do any code in here if condition is FALSE
  }
  
  ## Both option1 and option2 function identically. Some prerfer option2 considering it to easier to read. 
  
  ```
- The ifelse() function
  - Used for element-wise conditional testing
  - The function requies three arguemtns:
    - test: The conditional test
    - yes: what to do if TRUE
    - no:  what to do if FALSE
  - Syntax:
  ```
    ifelse(test = logical_expression, yes = True, no = False)
  ```
  
- The switch statement
  - Used to evaluate an epxression against a list of options
  - Syntax:
  ```
  switch (expression, list)
  
  ```
 
### Resources: R Conditional Statements & Functions
  - [If-Else in R](https://r-coder.com/if-else-r/)
  - [Logical Operators](https://stat.ethz.ch/R-manual/R-devel/library/base/html/Logic.html)
  - [Logical Operators](https://www.r-bloggers.com/2010/12/logical-operators-in-r/) by R-bloggers
  - [any()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/any) function to determine if some values are true
  - [all()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/all) function to determine if all of the values are true
  - [ifelse()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/ifelse) function provides conditional element selection
  - [switch()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/switch) statement to evaluate a condition against a list of options.

### Resources: Usefull R functions
  - [seq()](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/seq) used to generate an sequence of numbers

## Implement the R if statement
```{r Implement the R if statement, include=T, error=F, warning=F}






```

## Implement the any() and all() functions
```{r Implement the any() and all() functions, include=T, error=F, warning=F}





```



## Implement if-else logic in R
```{r Implement if-else logic in R, include=T, error=F, warning=F}





```
## Use the R ifelse() function, nesting, and the switch statement
```{r Use the R ifelse() function, nesting, and the switch statement, include=T, error=F, warning=F}

# The ifelse() function





# Nesting






# The switch statement







```
















